---
title: "Exercise 10"
author: "Nicole Navarro"
date: "November 14, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(dplyr)
library(tidyverse)
library(xlsx)
library(reshape2)
```


#Crime data

```{r, echo=TRUE}
#import data
filenames <- list.files("~/School/Fall 16/EDA/data/FDLE", 
                        pattern="*.xls", full.names=TRUE)

ldf <- lapply(filenames, function(x) 
  read.xlsx(x, sheetName ="Part II (1)", startRow=3))

CNames <- names(as.data.frame(ldf[1]))
Alachua_label <- lapply(ldf, function(x) 
  which((x[,1] == "Alachua County") | (x[,1] =="Alachua") ))
Washington_label <- lapply(ldf, function(x) 
  which((x[,1] == "Washington County") | (x[,1] == "Washington") ))

#check completeness
unlist(Washington_label)-unlist(Alachua_label)
```

```{r}
#get rid of col names (because they don't all match)
NULL_Name <- function(x) {
  names(x) <- NULL
  return(x)
}
ldf <- lapply(ldf, NULL_Name)
```

```{r}
#combine all years into one data frame
L<- length(as.data.frame(ldf[1])[1,])
FL_Crime_0612 <- as.data.frame(ldf[1])[Alachua_label: 
                                         Washington_label,]
for (i in 2:length(ldf)) {
  FL_Crime_0612 <- rbind(FL_Crime_0612, 
                         as.data.frame(ldf[i])[Alachua_label[[i]]: 
                                         Washington_label[[i]],1:L])
}

FL_Crime_0612 <- as.data.frame(FL_Crime_0612)
names(FL_Crime_0612) <-CNames
str(FL_Crime_0612)
```

```{r}
#clean up county names
FL_Crime_0612[,1] <- as.factor(gsub("\\ County", "", FL_Crime_0612[,1]))
FL_Crime_0612[,1] <- as.factor(gsub("Desoto", "DeSoto", FL_Crime_0612[,1]))

FL_Crime_0612[,2:length(FL_Crime_0612[1,])] <- sapply(FL_Crime_0612[,2:length(FL_Crime_0612[1,])], as.integer)

start <- rep(2006,67)
FL_Crime_0612$Year <- c(start,start+1,start+2,start+3,start+4,start+5,start+6)
colnames(FL_Crime_0612)[1] <- "County"
str(FL_Crime_0612)
```

```{r,eval=FALSE}
#save cleaned up version to csv
write.csv(FL_Crime_0612, "~/School/Fall 16/EDA/data/FDLE/FL_Crime_0612.csv")
```


#Educational Attainment Data

```{r}
#import data
filenames2 <- list.files("~/School/Fall 16/EDA/data/FLedu", pattern="*.csv", full.names=TRUE)

ldf2 <- lapply(filenames2, function(x) 
  read.csv(x, skip=1))


#combine into one frame
Cnames <- c("County","pop25up","HS25up","bachelors25up")
FL_edu_0912 <- as.data.frame(ldf2[1])[c(3,34,82,88)]
names(FL_edu_0912)<-Cnames
for (i in 2:length(ldf2)) {
  data <- as.data.frame(ldf2[i])[c(3,34,82,88)]
  names(data) <- Cnames
  FL_edu_0912 <- rbind(FL_edu_0912, data)
}

FL_edu_0912 <- as.data.frame(FL_edu_0912)
str(FL_edu_0912)

```


```{r}
#clean up county names
FL_edu_0912[,1] <- as.factor(gsub("\\ County, Florida", "", FL_edu_0912[,1]))
FL_edu_0912[,1] <- as.factor(gsub("Miami-Dade", "Miami Dade", FL_edu_0912[,1]))

FL_edu_0912[,2:length(FL_edu_0912[1,])] <- sapply(FL_edu_0912[,2:length(FL_edu_0912[1,])], as.numeric)


start <- rep(2009,67)
FL_edu_0912$Year <- c(start,start+1,start+2,start+3)
str(FL_edu_0912)
```

#Unemployment data

```{r}
#import data
filenames <- list.files("~/School/Fall 16/EDA/data/FLlabor", pattern="*.xls", full.names=TRUE)

#Read in only the data we want
ldf <- lapply(filenames, function(x) read.xlsx2(x, sheetIndex =1))
ldf <- lapply(ldf, function(x) x <- x[,c(4,5,10)])

CNames <- c("County","Year", "Unemployment_Rate") ##Future column names

Alachua_label <- lapply(ldf, function(x) grep("Alachua", x[,1])) ##starting lines in each frame
Washington_label <- lapply(ldf, function(x) grep("Washington County, FL", x[,1])) ##starting lines in each frame

ldf <- lapply(ldf, NULL_Name)  ##Naming tricks

#create data frame
FL_Unemployment_0612 <- as.data.frame(ldf[1])[Alachua_label[[1]]: 
                                         (Washington_label[[1]]),]

for (i in 2:length(ldf)) {
  FL_Unemployment_0612 <- rbind(FL_Unemployment_0612,
                         as.data.frame(ldf[i])[Alachua_label[[i]]:
                                         (Washington_label[[i]]),])
}
FL_Unemployment_0612 <- as.data.frame(FL_Unemployment_0612)
names(FL_Unemployment_0612) <-CNames
str(FL_Unemployment_0612)

```


```{r,echo=TRUE}
#clean up data frames

FL_Unemployment_0612[,1] <- as.factor(gsub("\\ County, FL", "", FL_Unemployment_0612[,1]))
FL_Unemployment_0612[,2] <- as.factor(gsub("\\ Year", "", FL_Unemployment_0612[,2]))
FL_Unemployment_0612[,1] <- as.factor(gsub("Miami-Dade", "Miami Dade", FL_Unemployment_0612[,1]))
FL_Unemployment_0612[,2] <- as.numeric(as.character(FL_Unemployment_0612[,2]))
FL_Unemployment_0612[,3] <- as.numeric(as.character(FL_Unemployment_0612[,3]))

rownames(FL_Unemployment_0612) <- NULL
str(FL_Unemployment_0612)


```

#Population Data

```{r}
#2010-2015
url <- "https://www.census.gov/popest/data/counties/totals/2015/files/CO-EST2015-alldata.csv"

popdata1015 <- read.csv(url, header=TRUE)
popFL1015 <- popdata1015 %>% select(c(6,7,10:15)) %>% filter(STNAME=="Florida")

#clean
popFL1015 <- popFL1015 %>% filter(CTYNAME != "Florida") %>% select(2:length(popFL1015[1,]))
head(popFL1015)

#change to numeric
popFL1015[,2:length(popFL1015[1,])] <- sapply(popFL1015[,2:length(popFL1015[1,])], as.character)
popFL1015[,2:length(popFL1015[1,])] <- sapply(popFL1015[,2:length(popFL1015[1,])], as.numeric)

#reshape (wide form to long form)
popFL1015_long <- melt(popFL1015, id.vars= "CTYNAME", measure.vars= c("POPESTIMATE2010", "POPESTIMATE2011", "POPESTIMATE2012", "POPESTIMATE2013", "POPESTIMATE2014", "POPESTIMATE2015"))
head(popFL1015_long)

#cleaning county names
names(popFL1015_long) <- c("County", "Year", "population")

popFL1015_long[,1] <- as.factor(gsub("\\ County", "", popFL1015_long[,1]))
popFL1015_long[,2] <- as.factor(gsub("[A-Z]", "", popFL1015_long[,2]))
popFL1015_long[,1] <- as.factor(gsub("Miami-Dade", "Miami Dade", popFL1015_long[,1]))

str(popFL1015_long)
```

```{r}
#2000-2009
url <- "https://www.census.gov/popest/data/intercensal/county/files/CO-EST00INT-TOT.csv"

popdata0009 <- read.csv(url, header=TRUE)
popFL0009 <- popdata0009 %>% select(c(6,7,9:18)) %>% filter(STNAME=="Florida")
popFL0009 <- popFL0009 %>% filter(CTYNAME != "Florida") %>% select(2:length(popFL0009[1,])) 

popFL0009[,2:length(popFL0009[1,])] <- sapply(popFL0009[,2:length(popFL0009[1,])], as.character)
popFL0009[,2:length(popFL0009[1,])] <-sapply(popFL0009[,2:length(popFL0009[1,])], as.numeric)

popFL0009_long <- melt(popFL0009, id.vars= "CTYNAME", measure.vars= c("POPESTIMATE2000", "POPESTIMATE2001", "POPESTIMATE2002", "POPESTIMATE2003", "POPESTIMATE2004", "POPESTIMATE2005", "POPESTIMATE2006", "POPESTIMATE2007", "POPESTIMATE2008", "POPESTIMATE2009"))

names(popFL0009_long) <- c("County", "Year", "population")

#clean up county names
popFL0009_long[,1] <- as.factor(gsub("\\ County", "", popFL0009_long[,1]))
popFL0009_long[,1] <- as.factor(gsub("Miami-Dade", "Miami Dade", popFL0009_long[,1]))
popFL0009_long[,2] <- as.factor(gsub("[A-Z]", "", popFL0009_long[,2]))
popFL0015 <- rbind(popFL0009_long, popFL1015_long)
popFL0015$Year <- as.integer(as.character(popFL0015$Year))

str(popFL0015)
```

```{r,eval=FALSE}
write.csv(popFL0015,"~/School/Fall 16/EDA/data/FLpopulation.csv")
```

```{r,message=FALSE}
#crime rates
FCdata <- inner_join(FL_Crime_0612, popFL0015)

FCdata <- FCdata %>% mutate(Assault_Rate = 10000*Simple.Assault/population) %>% mutate(Manslaughter_Rate = 10000*Manslaughter/population) %>% mutate(Drug_Rate = 10000*Drug.Arrest/population) %>% mutate(Embezzlement_Rate = 10000*Embezzlement/population) %>% mutate(Fraud_Rate = 10000*Fraud/population)

str(FCdata)
```

```{r}
#join crime, edu, and unemployment dfs
big_frame <- full_join(FCdata,FL_edu_0912, by = c("County","Year"))
big_frame <- full_join(big_frame, FL_Unemployment_0612, by = c("County","Year"))
str(big_frame)
```



#now for the fun stuff!


##Distributions

Assault Rate:

```{r,tidy=TRUE,echo=FALSE,warning=FALSE}
ggplot(big_frame, aes(x=Assault_Rate)) + 
  geom_histogram(binwidth=1)
```

Education:

```{r,tidy=TRUE,echo=FALSE,warning=FALSE}
ggplot(big_frame, aes(x=HS25up)) + 
  geom_histogram(binwidth=1)

ggplot(big_frame, aes(x=bachelors25up)) +
  geom_histogram(binwidth=1)
```

Unemployment:

```{r,tidy=TRUE,echo=FALSE,}
ggplot(big_frame, aes(x=Unemployment_Rate)) + 
  geom_histogram(binwidth=.2)
```

#Boxplots

```{r,echo=FALSE}
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}
```

####Assault Rates
```{r,echo=FALSE}
outliers <- big_frame[which(big_frame$Year == 2006),][which(is_outlier(big_frame[which(big_frame$Year == 2006),]$Assault_Rate) == TRUE),][,c(1,12,14)]
for (i in 2007:2012){
  outliers <- rbind(outliers,big_frame[which(big_frame$Year == i),][which(is_outlier(big_frame[which(big_frame$Year == i),]$Assault_Rate) == TRUE),][,c(1,12,14)])
}

ggplot(data=big_frame, aes(x=as.factor(Year), y=Assault_Rate)) +
  geom_boxplot() +
  geom_point(data=outliers,aes(x=as.factor(Year),y=Assault_Rate,color = County))

```

####Educational Attainment
```{r,echo=FALSE,warning=FALSE}
small_frame <- big_frame[which(complete.cases(big_frame) == TRUE),]
outliers <- small_frame[which(small_frame$Year == 2006),][which(is_outlier(small_frame[which(small_frame$Year == 2006),]$HS25up) == TRUE),][,c(1,12,20)]
for (i in 2007:2012){
  outliers <- rbind(outliers,small_frame[which(small_frame$Year == i),][which(is_outlier(small_frame[which(small_frame$Year == i),]$HS25up) == TRUE),][,c(1,12,20)])
}

ggplot(data=big_frame, aes(x=as.factor(Year), y=HS25up)) + 
  geom_boxplot() +
  geom_point(data=outliers,aes(x=as.factor(Year),y=HS25up,color = County))

ggplot(data=big_frame, aes(x=as.factor(Year), y=bachelors25up)) + 
  geom_boxplot()
```

####Unemployment rates
```{r,echo=FALSE}
outliers <- big_frame[which(big_frame$Year == 2006),][which(is_outlier(big_frame[which(big_frame$Year == 2006),]$Unemployment_Rate) == TRUE),][,c(1,12,22)]
for (i in 2007:2012){
  outliers <- rbind(outliers,big_frame[which(big_frame$Year == i),][which(is_outlier(big_frame[which(big_frame$Year == i),]$Unemployment_Rate) == TRUE),][,c(1,12,22)])
}

ggplot(data=big_frame, aes(x=as.factor(Year), y=Unemployment_Rate)) + 
  geom_boxplot() +
  geom_point(data=outliers,aes(x=as.factor(Year),y=Unemployment_Rate,color = County))
```

####Fraud rates
```{r,echo=FALSE}
outliers <- big_frame[which(big_frame$Year == 2006),][which(is_outlier(big_frame[which(big_frame$Year == 2006),]$Fraud_Rate) == TRUE),][,c(1,12,18)]
for (i in 2007:2012){
  outliers <- rbind(outliers,big_frame[which(big_frame$Year == i),][which(is_outlier(big_frame[which(big_frame$Year == i),]$Fraud_Rate) == TRUE),][,c(1,12,18)])
}

ggplot(data=big_frame, aes(x=as.factor(Year), y=Fraud_Rate)) + 
  geom_boxplot() + 
  geom_point(data=outliers,aes(x=as.factor(Year),y=Fraud_Rate,color = County))
```


####Drug arrest rates
```{r,echo=FALSE}
outliers <- big_frame[which(big_frame$Year == 2006),][which(is_outlier(big_frame[which(big_frame$Year == 2006),]$Drug_Rate) == TRUE),][,c(1,12,16)]
for (i in 2007:2012){
  outliers <- rbind(outliers,big_frame[which(big_frame$Year == i),][which(is_outlier(big_frame[which(big_frame$Year == i),]$Drug_Rate) == TRUE),][,c(1,12,16)])
}

ggplot(data=big_frame, aes(x=as.factor(Year), y=Drug_Rate)) + 
  geom_boxplot() +
  geom_point(data=outliers,aes(x=as.factor(Year),y=Drug_Rate,color = County))
```

####Manslaughter rates
```{r,echo=FALSE}
outliers <- big_frame[which(big_frame$Manslaughter_Rate > .75),][,c(1,12,15)]


ggplot(data=big_frame, aes(x=as.factor(Year), y=Manslaughter_Rate)) + 
  geom_boxplot() +
  geom_point(data=outliers,aes(x=as.factor(Year),y=Manslaughter_Rate, color = County))
```

####Embezzlement rates
```{r,echo=FALSE}
outliers <- big_frame[which(big_frame$Embezzlement_Rate > 2),][,c(1,12,17)]

ggplot(data=big_frame, aes(x=as.factor(Year), y=Embezzlement_Rate)) + 
  geom_boxplot() +
  geom_point(data=outliers,aes(x=as.factor(Year),y=Embezzlement_Rate,color = County))
```


##unemployment vs education


```{r,echo=FALSE,warning=FALSE}
ggplot(data=big_frame, aes(x=Unemployment_Rate, y=bachelors25up)) + 
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
cor(big_frame$Unemployment_Rate,big_frame$bachelors25up,use="complete")
```

##Unemployment and drug arrests
```{r,echo=FALSE}
ggplot(data=big_frame, aes(x=Unemployment_Rate, y=Drug_Rate)) + 
  geom_point(aes(color=as.factor(Year)))

ggplot(data=big_frame, aes(x=Unemployment_Rate, y=Drug_Rate)) + 
  geom_point() +
  geom_smooth(method="lm")

cor(big_frame$Unemployment_Rate,big_frame$Drug_Rate,use="complete")
```

##Education and drug arrests
```{r,echo=FALSE,warning=FALSE}

ggplot(data=big_frame, aes(x=bachelors25up, y=Drug_Rate)) + 
  geom_point() +
  geom_smooth(method="lm")
```
```{r,warning=FALSE}
cor(big_frame$bachelors25up,big_frame$Drug_Rate,use="complete")
```



##education and assault
```{r,echo=FALSE,warning=FALSE}

ggplot(data=big_frame, aes(x=bachelors25up, y=Assault_Rate)) + 
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
cor(big_frame$bachelors25up,big_frame$Assault_Rate,use="complete")
```

##education and manslaughter

```{r,echo=FALSE,warning=FALSE}

ggplot(data=big_frame, aes(x=bachelors25up, y=Manslaughter_Rate)) + 
  geom_point() +
  geom_smooth(method="lm")
```

```{r,warning=FALSE}
cor(big_frame$bachelors25up,big_frame$Manslaughter_Rate,use="complete")
```

#education and fraud

```{r,echo=FALSE,warning=FALSE}
ggplot(data=big_frame, aes(x=bachelors25up, y=Fraud_Rate)) + 
  geom_point() +
  geom_smooth(method="lm")
```
```{r}
cor(big_frame$bachelors25up,big_frame$Fraud_Rate,use="complete")
```



